### Vladimir Nimchenko

INTRODUCTION:

For my data, I took a Amazon reviews file of with the columns review title and review text. The data comes from Kaggle. I shortened the reviews to 100 (got the file to 100 rows) for simplicity purposes.


Loading the needed libraries
```{r}
library(tidyverse)
library(tidytext)
library(textdata)    
library(ggplot2)

```


Read the data into a data frame
```{r}
# Read in the data from Github into a data frame
amazon_reviews <- read.csv("https://raw.githubusercontent.com/GitHub-Vlad/Data-Science-Projects/main/Sentiment%20Analysis/Amazon_Reviews.csv", header = TRUE)

#View the data
View(amazon_reviews)
```

Sentimental Analysis for Review Title
```{r}
loughran_sentiments <- get_sentiments("loughran")
# convert review titles to words
amazon_title_words<- amazon_reviews %>%
  select(review_title) %>%
  mutate(linenumber = row_number()) %>%
  unnest_tokens(output = word, input = review_title, token = "words", format = "text", to_lower = TRUE)

# calculate the number of positive and negative words in the review title
amazon_title_sentiment <- amazon_title_words %>%
  inner_join(loughran_sentiments) %>%
  count(sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

# print the number of positive and negative words in the review title 
head(amazon_title_sentiment,10)

#count the most common positive and negative words in the review title
amazon_title_common_word_count <- amazon_title_words %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

# print the number and name of the most common positive and negative words in the review title 
head(amazon_title_common_word_count,10)

# graphing the top 5 positive and negative words
amazon_title_common_word_count %>%
  group_by(sentiment) %>%
  slice_max(n, n = 5) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Sentiment",
       y = NULL)
```

Sentimental Analysis for Review Text
```{r}
# convert review text to words
amazon_text_words<- amazon_reviews %>%
  select(review_text) %>%
  mutate(linenumber = row_number()) %>%
  unnest_tokens(output = word, input = review_text, token = "words", format = "text", to_lower = TRUE)

# calculate the number of positive and negative words in the review text
amazon_text_sentiment <- amazon_text_words%>%
  inner_join(loughran_sentiments) %>%
  count(sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

# print the number of positive and negative words in the review text
head(amazon_text_sentiment,10)

#count the most common positive and negative words in the review text
amazon_text_common_word_count <- amazon_text_words %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

# print the number and name of the most common positive and negative words in the review text
head(amazon_text_common_word_count,10)

# graphing the top 5 positive and negative words
amazon_text_common_word_count %>%
  group_by(sentiment) %>%
  slice_max(n, n = 5) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Sentiment",
       y = NULL)
```


CONCLUSION:

After conducting sentiment analysis for both the review text and the review title, I observed that review title had 10 negative and 22 positive words where as the review text had 133 positive and 104 negative words. This was a result that the text had many more words than the title. 